Performance benefits of all-memory processing for large neuroimaging data

### 2019-08-31 paper.tex
added a very rough intro

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Paper Title*}

\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address}
}

\maketitle

@@

\section{Introduction}
\section{Methods}
\section{Results}
\section{Conclusion}
\section{Acknowledgement}
\bibliographystyle{IEEEtran} 
\bibliography{biblio}

>>>>

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage[binary-units]{siunitx}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\newcommand{\bigdata}{Big Data }

\newcommand{\valerie}[1]{\color{blue}\textbf{Note from Valerie}:
      #1 \color{black}}

\begin{document}

\title{Performance benefits of all-memory processing for large neuroimaging data}

\author{\IEEEauthorblockN{Val\'erie Hayot-Sasson$^1$, Shawn T Brown$^2$ and 
    Tristan Glatard$^1$
  }\\
  \IEEEauthorblockA{
    $^1$Department of Computer Science and Software Engineering, Concordia University, Montreal, Canada\\
    $^2$Montreal Neurological Institute, McGill University, Montreal, Canada
  }
}
\maketitle

@@

\section{Introduction}
Data transfer rates are a significant source of bottleneck in \bigdata
applications. As a result, many \bigdata frameworks have opted to reduce data
transfers using strategies such as data locality, in-memory computing and lazy
evaluation. Furthermore, kernel optimizations, such as the Page Cache, also
attempt to minimize the amount of data transfers that may arise during the
lifetime of an application through cache hits. It also lessens the burden of
data transfers by asynchronously flushing to desired storage device.

Adaptations made for the efficient processing of \bigdata applications are not 
limited to software. Infrastructure technology has also improved to increase the
bandwidth of slower storage devices, to a point at which cost-effective Solid 
State Drives (SSD) can emulate RAM-like bandwidths and exhibit a significantly 
larger capacity that standard DRAM. Moreover, such SSDs provide persistent
storage enabling all data to be stored in a RAM-like non-volatile filesystem.
\valerie{Perhaps introduce optane here/do related-work search}

Neuroimaging open-data initiatives have led to extensively large repositories of
publicly available data. Such initiatives include the BigBrain~\cite{BigBrain}, 
a one-of-a-kind $\SI{603}{\gibi\byte}$\valerie{verify this w/ Claude}
histological image of a 76 year-old~\valerie{from memory, so better double-check} healthy human brain
at 20$\mu$m resolution; the UK Biobank~\cite{ukbiobank}, a repository expected to
contain approximately $\SI{0.2}{\peta\byte}$ of data (including various magnetic
resonance (MR) imaging modalities) from 500,000 individuals living in the UK;
the Human Connectome Project~\cite{HCP}, a repository containing MR scans from
1200 healthy adults, which is expected to exceed $\SI{2}{\peta\byte}$ in size; 
and the Consortium for Reliability and Reproducibility (CoRR)~\cite{corr}, an
initiative which aggregates MR data from various laboratories around the world, 
of which 32 are currently available and make up about $\SI{937}{\giga\byte}$ of
data in total.

Due to storage limitations, only subsets of the neuroimaging repositories 
can be processed in a typical research laboratory. Moreover, as these datasets 
are extremely large and are only increasing in size, they are typically stored 
in higher-capacity, slower storage devices. In such conditions, large-scale 
studies in neuroimaging remain limited to labs with access to adequate 
infrastructure and are limited by data transfer times.

In this paper, we aim to:
\begin{itemize}
        \item Quantify the added value of Intel Optane persistent memory on 
            processing large neuroimaging data using representative pipelines; and
        \item Determine when a given Intel Optane storage configuration (Memory 
            and App Direct mode) is preferable.
\end{itemize}

\section{Materials and Methods}
\subsection{Infrastructure}
\subsection{Storage configuration}
\subsection{Applications}
\subsection{Performance model}
\section{Results}
\subsection{Memory-mode}
\subsection{App-Direct mode}
\section{Discussion}
\subsection{Memory vs App Direct mode}
\subsection{Effect of Page Cache}
\subsection{Added value of persistent memory}
\section{Conclusion}
\section{Acknowledgement}
\bibliographystyle{IEEEtran} 
\bibliography{biblio}
